The Strong Gravitational Lens Finding Challenge

### 2018-01-29 paper.tex
Edit.

\title[SL Challenge]{The Strong Gravitational  Lens Finding Challenge}
%C.~E.~Petrillo$^{1}$,  C.~Tortora$^{1}$, L.~V.~E.~Koopmans$^{1}$, \and G.~Verdoes~Kleijn$^{1}$
\author[R. B. Metcalf {\it et al.}]{R. Benton Metcalf,$^{1,4}$\thanks{E-mail: robertbenton.metcalf@unibo.it} 
M. Meneghetti,$^4$ 
Camille Avestruz,$^{11,12,13,\dagger}$
\newauthor
Fabio Bellagamba,$^1$
Cl\'ecio R. Bom,$^{17,18}$
Emmanuel Bertin,$^3$ 
\newauthor
R\'emi Cabanac,$^2$ 
Etienne Decenci\`ere,$^{10}$
R\'emi Flamary,$^8$
\newauthor
Raphael Gavazzi,$^{3}$ 
Mario Geiger,$^6$ 
Philippa Hartley,$^7$
\newauthor
Marc Huertas-Company,$^9$
Neal Jackson,$^7$ 
Eric Jullo,$^{16}$
\newauthor
Jean-Paul Kneib,$^6$
L\'{e}on V. E. Koopmans,$^{21}$
\newauthor
Fran\c{c}ois Lanusse,$^{19}$ 
Chun-Liang Li,$^{20}$
Quanbin Ma,$^{20}$ 
\newauthor
Martin Makler,$^{18}$
Nan Li,$^{14}$
Matthew Lightman,$^{15}$
\newauthor
Carlo Enrico Petrillo,$^{21}$
Christoph Sch\"{a}fer,$^6$ 
Alessandro Sonnenfeld,$^5$ 
\newauthor
Amit Tagore,$^7$
Crescenzo Tortora,$^{21}$
Diego Tuccillo,$^{9,10}$
\newauthor
Manuel B. Valent\'in,$^{18}$
Santiago Velasco-Forero,$^{10}$
Gijs A. Verdoes Kleijn,$^{21}$
\newauthor
and Georgios Vernardos$^{21}$
\\
% List of institutions
%\vspace{0.5cm}\\
$^{1}$ Departimento di Fisica \& Astronomia, Universit\`a di Bologna, via Gabetti 93/2, 4012 Bologna, Italy \\
$^{2}$ IRAP, Universit\'e de Toulouse, CNRS, UPS, Toulouse, France.\\
$^{3}$ Sorbonne Universit\'{e}s, UPMC Univ. Paris 6 and CNRS, UMR 7095, Institut d'Astrophysique de Paris,  \\ 98bis Bd Arago, F-75014, Paris France.\\
$^{4}$ INAF-Osservatorio Astronomico di Bologna, via Ranzani 1, 40127 Bologna, Italy \\
$^{5}$ Kavli IPMU (WPI), UTIAS, The University of Tokyo, Kashiwa, Chiba 277-8583, Japan\\
$^{6}$ Institute of Physics, Laboratory of Astrophysics, Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), \\ Observatoire de Sauverny, 1290 Versoix, Switzerland \\
$^{7}$ Jodrell Bank Centre for Astrophysics, School of Physics \& Astronomy, University of \\ Manchester, Oxford Rd, Manchester M13 9PL, UK \\
$^{8}$ Laboratoire Lagrange, Universi\'{e} de Nice Sophia-Antipolis, Centre National de la Recherche Scientifique,\\ Observatoire de la C\^{o}te d'Azur, Parc Valrose, 06108 Nice, France \\
$^{9}$ LERMA, Observatoire de Paris, CNRS, Universit\'e Paris Diderot, 61, Avenue de l'Observatoire F-75014, Paris, France \\
$^{10}$ MINES Paristech, PSL Research University, Centre for Mathematical Morphology, 35 rue \\ Saint-Honore, Fontainebleau, France \\
$^{11}${Enrico Fermi Institute, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^{12}${Kavli Institute for Cosmological Physics, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^{13}${Department of Astronomy \& Astrophysics, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^{14}${School of Physics and Astronomy, Nottingham University, University Park, Nottingham, NG7 2RD, UK
}\\
$^{15}${JPMorgan Chase, Chicago, IL 60603 U.S.A.}\\
$^{16}$Aix Marseille Universit\'e, CNRS, LAM (Laboratoire d'Astrophysique de Marseille) UMR 7326, 13388, Marseille, France \\
$^{17}${Centro Federal de Educa\c{c}\~ao Tecnol\'ogica Celso Suckow da Fonseca, CEP 23810-000,  Itagua\'i, RJ, Brazil}\\
$^{18}${Centro Brasileiro de Pesquisas F\'isicas, CEP 22290-180, Rio de Janeiro, RJ, Brazil}\\
$^{19}${McWilliams Center for Cosmology, Department of Physics, Carnegie Mellon University, Pittsburgh, PA 15213, USA}\\
$^{20}${School of Computer Science, Carnegie Mellon University, Pittsburgh, PA 15213, USA}\\
$^{21}${Kapteyn Astronomical Institute, University of Groningen, Postbus 800, 9700 AV, Groningen, The Netherlands}\\
$^\dagger${Provost's Postdoctoral Scholar at the University of Chicago}\\
}

@@

% Abstract of the paper
\begin{abstract}
Large scale imaging surveys will increase the number of galaxy-scale strong lensing candidates by maybe three orders of magnitudes beyond the number known today.  Finding these rare objects will require picking them out of at least tens of millions of images and deriving scientific results from them will require quantifying the efficiency and bias of any search method.  To achieve these objectives automated methods must be developed.
Because gravitational lenses are rare objects reducing false positives will be particularly important.  We present a description and results of an open gravitational lens finding challenge.  Participants were asked to classify 100,000 candidate objects as to whether they were gravitational lenses or not with the goal of developing better automated methods for finding lenses in large data sets.   A variety of methods were used including visual inspection, arc and ring finders, support vector machines (SVM) and convolutional neural networks (CNN).  We find that many of the methods will be easily fast enough to analyze the anticipated data flow.  In test data several methods are able to identify upwards of half the lenses after applying some thresholds on the lens characteristics such as lensed image brightness, size or contrast with the lens galaxy without making a single false-positive identification.   This is significantly better than direct inspection by humans was able to do.  Having multi-band, ground based data is found to be better for this purpose than single-band space based data with lower noise and higher resolution.
The most difficult challenge for a lens finder is differentiating between rare irregular and ring galaxies and true gravitational lenses.  The degree to which the efficiency and biases of lens finders can be quantified largely depends on the realism of the simulated data on which the finders are trained.
\end{abstract}

@@

Up to this point, the most widely used method for finding lenses in imaging surveys has been by visual inspection of candidates that have been selected on the bases of luminosity and/or color.  This has been done in the radio \citep{2003MNRAS.341...13B} and in the visible with space and ground based data \citep{2008MNRAS.389.1311J,2008ApJS..176...19F,2014MNRAS.439.3392P}.  A related method pioneered by the SPACE WARPS project \citep{2016MNRAS.455.1171M,2016MNRAS.455.1191M,2015MNRAS.452..502G} has been to crowd source the visual inspection through an online platform.  These efforts have proved very fruitful, but they deal with orders of magnitude fewer candidates and lenses than will be necessary in the future.  Dealing with such large quantities of data will not be practical for any visual inspection approach.  In addition, the efficiency and detection bias of human inspection methods are difficult to rigorously quantify.

Spectroscopic searches for galaxy scale lenses have also been done by looking for high redshift stellar lines in the spectra of lower redshift large galaxies.  Notably this was done in the Sloan Lens ACS (SLACS) survey, producing a relatively well defined and pure sample of Einstein ring lenses \citep{2006ApJ...638..703B}.  However, spectroscopy is telescope-time consuming and for the foreseeable future we are not likely to have spectroscopic surveys that cover anywhere near the number of objects as the planned imaging surveys.

Some automated algorithms have been developed in the past to detect lenses in images.  These have been designed to detect arc-like features \citep{2006astro.ph..6757A,2007A&A...472..341S, 2017A&A...597A.135B} and rings \citep{2014ApJ...785..144G,2014A&A...566A..63J}.  They have been applied to survey data and found of order 200 lenses \citep{2007A&A...461..813C,2012ApJ...749...38M,2016A&A...592A..75P}.  Another approach has been to use a lens modeling code to fit a model to all the candidates and pick out the ones that fit the model well \citep{2009ApJ...694..924M,2010A&A...517A..25S,2017arXiv170401585S}.

@@

The paper is organized as follows.  The form of the challenge and its rules are described in the next section.  The methods used to simulate mock images of galaxies and gravitational lenses are described in section~\ref{sec:simulation}.  In section~\ref{sec:entries}, each of the methods that were used to solve the challenge are briefly described.  We discuss the metrics used to evaluate entries in section~\ref{sec:figure_of_merit}.  The performance of each of the methods broken down in different ways is presented in section~\ref{sec:performance}.  Finally,  in section~\ref{sec:conclusion}, we conclude with a discussion of what was learned and how methods can be improved in the future.

@@

Creating the mock images started with a cosmological N-body simulation.  In this case it was the Millennium simulation \citep{2009MNRAS.398.1150B}.   A catalog of dark matter halos within a light-cone is constructed within the Millennium Observatory project \citep{2013MNRAS.428..778O}.   The challenge sets were based on a 1.6 sq.deg. light cone extending out to redshift $z=6$.  The halos are found with a friends-of-friends algorithm and characterized by a total mass, size and half mass radius.  They include subhalos of larger halos.  The halo catalog is read into the GLAMER lensing code \citep{2014MNRAS.445.1942M,2014MNRAS.445.1954P} to do all the ray-tracing.    Within the code a Navarro, Frenk \& White (NFW) \citep{1996ApJ...462..563N} profile is fit to the three parameters given above.  The halos are projected onto a series of lens planes, in this case 20,  and the deflection angle at any point in each plane can be calculated by summing the effects of all the halos with a hybrid tree method.  In this way the halos have the mass, concentration and clustering properties from the N-body simulation, but within each strong lens the mass resolution is not limited by the original simulation, but follows the analytic mass profile.   An additional mass component that will be discussed later is added to each halo to represent the stellar ans gas mass.

@@

The furthest point in the caustic is found from its own center and the source is placed randomly within 3 times this distance.  This is a somewhat arbitrary length designed to be a compromise between producing only very clear strong lenses because all the sources are right in the center of the caustic and making the process  inefficient because most of the sources are too far away from the caustic to produce clear lenses.  If the source positions where taken completely at random the fraction of clear lenses would be very low.

The visible galaxies associated with the lens must also be simulated. There  are too few bright galaxies in the UDF catalog  to make enough mock lens galaxies for this purpose.  Instead, for most of the lenses, we used an analytic model for the surface brightness of these galaxies.  The Millennium Observatory provides parameters for the galaxies that inhabit the dark matter halos using the semi-analytic galaxy formation models of \citet{2011MNRAS.413..101G}.  The parameters used here were the total magnitude, the bulge-to-disk ratio, the disk scale height and the bulge effective radius.  The magnitude and bulge-to-disk ratio are a function of the pass band.  Each galaxy is given a random orientation and inclination angle.between 0 and $80\degree$.  The disk is exponential with no vertical height which is why the inclination is limited to $80\degree$.  The bulge is represented by a S\'{e}rsic profile with a axis ratio randomly sampled between 0.5 and 1.  The S\'{e}rsic index, $n_s$, is given by 
\begin{align}
\log( n_s )= 0.4 \log\left[ {\rm max}\left(\frac{B}{T},0.03\right)\right] + 0.1 x 
\end{align}
where $\frac{B}{T}$ is the bulge to total flux ratio and $x$ is a uniform random number between -1 and 1.  This approximately reproduces the observed correlation between these quantities.

@@

To represent the mass of the galaxies we make a gridded map of the surface brightness at 3 times the resolution of the final image.  The surface brightness map in converted into a mass map withing GLAMER by assuming a uniform mass-to-light ratio of 1.5 solar in the reference band.  These mass maps are added to the NFW dark matter halos discussed before to make the total lens mass distribution.  The deflections caused by the mass maps are calculated by Fast Fourier Transform (FFT) and added to the halos'  deflections for the ray tracing is.

The code is able to produce any combination of foreground galaxies, lensed image and noise that is desired.  For the training set we provided an image of the total lens with noise, an image of the foreground galaxies with noise and image of the lensed background source without noise.
For the test sets only the final images were provided to participants although all the information was stored for analyzing the entries.

@@

By chance one of the original KiDS images appears to have been a lens.  When an additional lensed source was added this made a double lens or "jackpot" lens.

@@

\begin{table*}
\centering
\begin{tabular}{rlll}
  \hline
 & Name & type & authors \\ 
  \hline
  1 & AstrOmatic & Space-Based & Bertin   \\ 
  2 & GAHEC IRAP & Space-Based & Cabanac  \\ 
  3 & \red{CAS Swinburne Melb} & Ground-Based & Jacobs   \\ 
  4 & ALL-star & Ground-Based & Avestruz, N.Li \& Lightman  \\ 
  5 & Manchester1 & Space-Based & Jackson \& Tagore  \\ 
  6 & CMU-DeepLens-Resnet-Voting & Space-Based & Ma, Lanusse \& C. Li  \\ 
  7 & Manchester SVM & Ground-Based & Hartley \& Flamary  \\ 
  8 & CMU-DeepLens-ResNet & Space-Based & Francois Lanusse, Ma, C. Li \& Ravanbakhsh \\ 
  9 & CMU-DeepLens-Resnet-Voting & Ground-Based & Ma, Lanusse \& C. Li   \\ 
  10 & YattaLensLite & Space-Based &  Sonnenfeld  \\ 
  11 & \red{NeuralNet2} & Space-Based &  Davies  \\ 
  12 & CAST & Ground-Based &  Roque De Bom, Valent\'{\i}n \&  Makler  \\ 
  13 & CMU-DeepLens-ResNet-ground3 & Ground-Based &  Lanusse, Ma, Ravanbakhsh \& C. Li \\ 
  14 & GAMOCLASS & Space-Based & Marc Huertas-Company, Tuccillo, Velasco-Forero \& Decenci\~{a}re \\ 
  15 & LASTRO EPFL (CNN) & Space-Based & Geiger, Sch\"{a}fer \& Kneib \\ 
  16 & Manchester SVM & Space-Based &  Hartley  \& Flamary\\ 
  17 & CMU-DeepLens-ResNet-aug & Space-Based & Ma,  Lanusse, Ravanbakhsh \& C. Li \\ 
  18 & LASTRO EPFL & Ground-Based & Geiger, Sch\"{a}fer \& Kneib   \\ 
  19 & CAST & Space-Based & Bom, Valent\'{\i}n \& Makler  \\ 
  20 & AstrOmatic & Ground-Based & Bertin  \\ 
  21 & All-now & Space-Based & Avestruz, N. Li \& Lightman   \\ 
  22 & Manchester-NA2 & Ground-Based & Jackson \& Tagore  \\ 
  23 & YattaLensLite & Ground-Based & Sonnenfeld   \\ 
  24 & Kapteyn Resnet & Space-Based & Petrillo, Tortora, Verdoes Kleijn, Koopmans \& Vernardos \\ 
   \hline
\end{tabular}
\caption{Entries to the challenges.}
\label{table:entries}
\end{table*}

@@

Arcfinder \citep{2006astro.ph..6757A,2007A&A...461..813C,2012ApJ...749...38M} illustrated in Figure \ref{fig:Cabanac}, is a fast linear method that computes a pixel wise elongation parameter (ratio of first-order moments in a n-pix window oriented in proper reference frame) for all pixels of mexican-hat-smoothed FITS images. Arcfinder then extracts contiguous pixels above a given background and computes the candidate arc's length, width, area, radius of curvature and peak surface brightness. A final thresholding is set to maximize purity over completeness on a few typical arcs of the dataset.
For the current SL challenge, arcfinder was tuned to detect long and narrow arcs, and was optimized on a subset of 1000 simulated images with a grid covering a range of elongation windows and arc areas.  A python wrapper allows users to change parameters in a flexible way and run the arcfinder C code as a linux line command. Arcfinder took a couple of hours to run on the entire dataset with some overheads due to the dataset format. The code is publicly available at https://github.com/rcabanac/arcfinder.

@@

For a binary classification problem such as the one involved in strong lens detection, the training is performed by optimizing the weights of the model as to minimize the \textit{binary cross-entropy} objective:
\begin{equation}
	- \sum_{n=1}^{N} y[n] \log \hat{y}[n] + (1 - y[n])\log(1 - \hat{y}[n]) \;,
\end{equation}
where $N$ is the number of training instances, $y \in \{0,1\}$ is the true class of the image and $\hat{y} \in [0,1]$ is the class probability predicted by the model. This optimization is usually performed by a Stochastic Gradient Descent (SGD) algorithm or its variants (e.g. ADAM \citep{Kingma_2014}, Adagrad \citep{duchi12adaptive}, RMSprop \citep{Tieleman12RMSProp}, or accelerated gradients \citep{nesterov83method}). SGD updates the model iteratively by taking small gradient steps over randomly selected subsamples of the training set (so called \textit{mini-batches}). All the CNN-based methods presented in this work rely on the ADAM optimization algorithm, which also uses past gradients from previous iterations to adaptively estimate lower-order moments. Empirically it has been found that in many problems ADAM converges faster than SGD \citep{Ruder16overview}.

To avoid overfitting in neural networks, we used the Dropout regularization technique \citep{2012arXiv1207.0580H,JMLR:v15:srivastava14a}. During training, we randomly drop a certain percent of neurons and their connections from the neural
network. This regularization technique reduces overfitting  by preventing complex co-adaptations of neurons on training data.

>>>>

\title[SL Challenge]{The Strong Gravitational  Lens Finding Challenge}
%C.~E.~Petrillo$^{1}$,  C.~Tortora$^{1}$, L.~V.~E.~Koopmans$^{1}$, \and G.~Verdoes~Kleijn$^{1}$
\author[R. B. Metcalf {\it et al.}]{R. Benton Metcalf,$^{1,4}$\thanks{E-mail: robertbenton.metcalf@unibo.it} 
M. Meneghetti,$^4$ 
Camille Avestruz,$^{11,12,13,\dagger}$
\newauthor
Fabio Bellagamba,$^1$
Cl\'ecio R. Bom,$^{17,18}$
Emmanuel Bertin,$^3$ 
\newauthor
R\'emi Cabanac,$^2$ 
Etienne Decenci\`ere,$^{10}$
R\'emi Flamary,$^8$
\newauthor
Raphael Gavazzi,$^{3}$ 
Mario Geiger,$^6$ 
Philippa Hartley,$^7$
\newauthor
Marc Huertas-Company,$^9$
Neal Jackson,$^7$ 
Eric Jullo,$^{16}$
\newauthor
Jean-Paul Kneib,$^6$
L\'{e}on V. E. Koopmans,$^{21}$
\newauthor
Fran\c{c}ois Lanusse,$^{19}$ 
Chun-Liang Li,$^{20}$
Quanbin Ma,$^{20}$ 
\newauthor
Martin Makler,$^{18}$
Nan Li,$^{14}$
Matthew Lightman,$^{15}$
\newauthor
Carlo Enrico Petrillo,$^{21}$
Christoph Sch\"{a}fer,$^6$ 
Alessandro Sonnenfeld,$^5$ 
\newauthor
Amit Tagore,$^7$
Crescenzo Tortora,$^{21}$
Diego Tuccillo,$^{9,10}$
\newauthor
Manuel B. Valent\'in,$^{18}$
Santiago Velasco-Forero,$^{10}$
Gijs A. Verdoes Kleijn,$^{21}$
\newauthor
and Georgios Vernardos$^{21}$
\\
% List of institutions
%\vspace{0.5cm}\\
$^{1}$ Departimento di Fisica \& Astronomia, Universit\`a di Bologna, via Gabetti 93/2, 4012 Bologna, Italy \\
$^{2}$ IRAP, Universit\'e de Toulouse, CNRS, UPS, Toulouse, France.\\
$^{3}$ Institut d'Astrophysique de Paris, Sorbonne Universit\'e, CNRS,  UMR 7095, 98 bis bd Arago, 75014 Paris, France.\\
$^{4}$ INAF-Osservatorio Astronomico di Bologna, via Ranzani 1, 40127 Bologna, Italy \\
$^{5}$ Kavli IPMU (WPI), UTIAS, The University of Tokyo, Kashiwa, Chiba 277-8583, Japan\\
$^{6}$ Institute of Physics, Laboratory of Astrophysics, Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), \\ Observatoire de Sauverny, 1290 Versoix, Switzerland \\
$^{7}$ Jodrell Bank Centre for Astrophysics, School of Physics \& Astronomy, University of \\ Manchester, Oxford Rd, Manchester M13 9PL, UK \\
$^{8}$ Laboratoire Lagrange, Universi\'{e} de Nice Sophia-Antipolis, Centre National de la Recherche Scientifique,\\ Observatoire de la C\^{o}te d'Azur, Parc Valrose, 06108 Nice, France \\
$^{9}$ LERMA, Observatoire de Paris, CNRS, Universit\'e Paris Diderot, 61, Avenue de l'Observatoire F-75014, Paris, France \\
$^{10}$ MINES Paristech, PSL Research University, Centre for Mathematical Morphology, 35 rue \\ Saint-Honore, Fontainebleau, France \\
$^{11}${Enrico Fermi Institute, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^{12}${Kavli Institute for Cosmological Physics, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^{13}${Department of Astronomy \& Astrophysics, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^{14}${School of Physics and Astronomy, Nottingham University, University Park, Nottingham, NG7 2RD, UK
}\\
$^{15}${JPMorgan Chase, Chicago, IL 60603 U.S.A.}\\
$^{16}$Aix Marseille Universit\'e, CNRS, LAM (Laboratoire d'Astrophysique de Marseille) UMR 7326, 13388, Marseille, France \\
$^{17}${Centro Federal de Educa\c{c}\~ao Tecnol\'ogica Celso Suckow da Fonseca, CEP 23810-000,  Itagua\'i, RJ, Brazil}\\
$^{18}${Centro Brasileiro de Pesquisas F\'isicas, CEP 22290-180, Rio de Janeiro, RJ, Brazil}\\
$^{19}${McWilliams Center for Cosmology, Department of Physics, Carnegie Mellon University, Pittsburgh, PA 15213, USA}\\
$^{20}${School of Computer Science, Carnegie Mellon University, Pittsburgh, PA 15213, USA}\\
$^{21}${Kapteyn Astronomical Institute, University of Groningen, Postbus 800, 9700 AV, Groningen, The Netherlands}\\
$^\dagger${Provost's Postdoctoral Scholar at the University of Chicago}\\
}

@@

% Abstract of the paper
\begin{abstract}
Large scale imaging surveys will increase the number of galaxy-scale strong lensing candidates by maybe three orders of magnitudes beyond the number known today.  Finding these rare objects will require picking them out of at least tens of millions of images and deriving scientific results from them will require quantifying the efficiency and bias of any search method.  To achieve these objectives automated methods must be developed.
Because gravitational lenses are rare objects reducing false positives will be particularly important.  We present a description and results of an open gravitational lens finding challenge.  Participants were asked to classify 100,000 candidate objects as to whether they were gravitational lenses or not with the goal of developing better automated methods for finding lenses in large data sets.   A variety of methods were used including visual inspection, arc and ring finders, support vector machines (SVM) and convolutional neural networks (CNN).  We find that many of the methods will be easily fast enough to analyse the anticipated data flow.  In test data, several methods are able to identify upwards of half the lenses after applying some thresholds on the lens characteristics such as lensed image brightness, size or contrast with the lens galaxy without making a single false-positive identification.  This is significantly better than direct inspection by humans was able to do.  Having multi-band, ground based data is found to be better for this purpose than single-band space based data with lower noise and higher resolution, suggesting that colours bring a crucial additional information.
The most difficult challenge for a lens finder is differentiating between rare irregular and ring-like face-on galaxies and true gravitational lenses.  The degree to which the efficiency and biases of lens finders can be quantified largely depends on the realism of the simulated data on which the finders are trained.
\end{abstract}

@@

Up to this point, the most widely used method for finding lenses in imaging surveys has been by visual inspection of candidates that have been selected on the bases of luminosity and/or colour.  This has been done in the radio \citep{2003MNRAS.341...13B} and in the visible with space and ground based data \citep{2008MNRAS.389.1311J,2008ApJS..176...19F,2010A&A...517A..25S,2014MNRAS.439.3392P}.  A related method pioneered by the SPACE WARPS project \citep{2016MNRAS.455.1171M,2016MNRAS.455.1191M,2015MNRAS.452..502G} has been to crowd source the visual inspection through an online platform.  These efforts have proved very fruitful, but they deal with orders of magnitude fewer candidates and lenses than will be necessary in the future.  Dealing with such large quantities of data will not be practical for any visual inspection approach.  In addition, the efficiency and detection bias of human inspection methods are difficult to rigorously quantify.

Spectroscopic searches for galaxy scale lenses have also been done by looking for high redshift stellar lines in the spectra of lower redshift large galaxies.  Notably this was done in the Sloan Lens ACS (SLACS) survey, producing a relatively well defined and pure sample of Einstein ring lenses \citep{2006MNRAS.369.1521W,2006ApJ...638..703B,2012ApJ...744...41B,2015MNRAS.449.3441S}.  However, spectroscopy is telescope-time consuming and for the foreseeable future we are not likely to have spectroscopic surveys that cover anywhere near the number of objects as the planned imaging surveys.

Some automated algorithms have been developed in the past to detect lenses in images.  These have been designed to detect arc-like features \citep{2006astro.ph..6757A,2007A&A...472..341S, 2017A&A...597A.135B} and rings \citep{2014ApJ...785..144G,2014A&A...566A..63J}.  They have been applied to survey data and found of order 200 lenses \citep{2007A&A...461..813C,2012ApJ...749...38M,2016A&A...592A..75P}.  Another approach has been to use a lens modeling code to fit a model to all the candidates and pick out the ones that fit the model well \citep{2009ApJ...694..924M,2017arXiv170401585S}.

@@

The paper is organised as follows.  The form of the challenge and its rules are described in the next section.  The methods used to simulate mock images of galaxies and gravitational lenses are described in section~\ref{sec:simulation}.  In section~\ref{sec:entries}, each of the methods that were used to solve the challenge are briefly described.  We discuss the metrics used to evaluate entries in section~\ref{sec:figure_of_merit}.  The performance of each of the methods broken down in different ways is presented in section~\ref{sec:performance}.  Finally,  in section~\ref{sec:conclusion}, we conclude with a discussion of what was learned and how methods can be improved in the future.

@@

Creating the mock images started with a cosmological N-body simulation.  In this case it was the Millennium simulation \citep{2009MNRAS.398.1150B}.   A catalog of dark matter halos within a light-cone is constructed within the Millennium Observatory project \citep{2013MNRAS.428..778O}.   The challenge sets were based on a 1.6 sq.deg. light cone extending out to redshift $z=6$.  The halos are found with a friends-of-friends algorithm and characterised by a total mass, size and half mass radius.  They include subhalos of larger halos.  The halo catalog is read into the GLAMER lensing code \citep{2014MNRAS.445.1942M,2014MNRAS.445.1954P} to do all the ray-tracing.    Within the code a Navarro, Frenk \& White (NFW) \citep{1996ApJ...462..563N} profile is fit to the three parameters given above.  The halos are projected onto a series of lens planes, in this case 20,  and the deflection angle at any point in each plane can be calculated by summing the effects of all the halos with a hybrid tree method.  In this way the halos have the mass, concentration and clustering properties from the N-body simulation, but within each strong lens the mass resolution is not limited by the original simulation, but follows the analytic mass profile.   An additional mass component that will be discussed later is added to each halo to represent the stellar and gas mass \red{I don't see any further reference to the gas component}.

@@

The furthest point in the caustic is found from its own center and the source is placed randomly within 3 times this distance.  This is a somewhat arbitrary length designed to be a compromise between producing only very clear strong lenses because all the sources are right in the center of the caustic and making the process inefficient because most of the sources are too far away from the caustic to produce clear lenses.  If the source positions were taken completely at random the fraction of clear lenses would be very low.

The visible galaxies associated with the lens must also be simulated. There  are too few bright galaxies in the UDF catalog  to make enough mock lens galaxies for this purpose.  Instead, for most of the lenses, we used an analytic model for the surface brightness of these galaxies.  The Millennium Observatory provides parameters for the galaxies that inhabit the dark matter halos using the semi-analytic galaxy formation models of \citet{2011MNRAS.413..101G}.  The parameters used here were the total magnitude, the bulge-to-disk ratio, the disk scale height and the bulge effective radius.  The magnitude and bulge-to-disk ratio are a function of the pass band.  Each galaxy is given a random orientation and inclination angle between 0 and $80\degree$.  The disk is exponential with no vertical height which is why the inclination is limited to $80\degree$.  The bulge is represented by a S\'{e}rsic profile with an axis ratio randomly sampled between 0.5 and 1.  The S\'{e}rsic index, $n_s$, is given by 
\begin{align}
\log( n_s )= 0.4 \log\left[ {\rm max}\left(\frac{B}{T},0.03\right)\right] + 0.1 x 
\end{align}
where $\frac{B}{T}$ is the bulge to total flux ratio and $x$ is a uniform random number between -1 and 1.  This approximately reproduces the observed correlation between these quantities \red{REF?}.

@@

To represent the mass of the galaxies we make a gridded map of the surface brightness at 3 times the resolution of the final image.  The surface brightness map in converted into a mass map within GLAMER by assuming a uniform mass-to-light ratio of 1.5 solar in the reference band.  These mass maps are added to the NFW dark matter halos discussed before to make the total lens mass distribution.  The deflections caused by the mass maps are calculated by Fast Fourier Transform (FFT) and added to the halos'  deflections for the ray tracing.

The code is able to produce any combination of foreground galaxies, lensed image and noise that is desired.  For the training set we provided an image of the total lens with noise, an image of the foreground galaxies with noise and image of the lensed background source without noise.
For the test sets only the final images were provided to participants although all the information was stored for analysing the entries.

@@

By chance one of the original KiDS images appears to have been a lens.  When an additional lensed source was added this made a double lens or "jackpot" lens \citep{2008ApJ...677.1046G}.

@@

\begin{table*}
\centering
\begin{tabular}{rlll}
  \hline
 & Name & type & authors \\ 
  \hline
  1 & AstrOmatic & Space-Based & Bertin   \\ 
  2 & GAHEC IRAP & Space-Based & Cabanac  \\ 
  3 & \red{CAS Swinburne Melb} & Ground-Based & Jacobs   \\ 
  4 & ALL-star & Ground-Based & Avestruz, N. Li \& Lightman  \\ 
  5 & Manchester1 & Space-Based & Jackson \& Tagore  \\ 
  6 & CMU-DeepLens-Resnet-Voting & Space-Based & Ma, Lanusse \& C. Li  \\ 
  7 & Manchester SVM & Ground-Based & Hartley \& Flamary  \\ 
  8 & CMU-DeepLens-ResNet & Space-Based & Francois Lanusse, Ma, C. Li \& Ravanbakhsh \\ 
  9 & CMU-DeepLens-Resnet-Voting & Ground-Based & Ma, Lanusse \& C. Li   \\ 
  10 & YattaLensLite & Space-Based &  Sonnenfeld  \\ 
  11 & \red{NeuralNet2} & Space-Based &  Davies  \\ 
  12 & CAST & Ground-Based &  Roque De Bom, Valent\'{\i}n \&  Makler  \\ 
  13 & CMU-DeepLens-ResNet-ground3 & Ground-Based &  Lanusse, Ma, Ravanbakhsh \& C. Li \\ 
  14 & GAMOCLASS & Space-Based &  Huertas-Company, Tuccillo, Velasco-Forero \& Decenci\`{e}re \\ 
  15 & LASTRO EPFL (CNN) & Space-Based & Geiger, Sch\"{a}fer \& Kneib \\ 
  16 & Manchester SVM & Space-Based &  Hartley  \& Flamary\\ 
  17 & CMU-DeepLens-ResNet-aug & Space-Based & Ma,  Lanusse, Ravanbakhsh \& C. Li \\ 
  18 & LASTRO EPFL & Ground-Based & Geiger, Sch\"{a}fer \& Kneib   \\ 
  19 & CAST & Space-Based & Bom, Valent\'{\i}n \& Makler  \\ 
  20 & AstrOmatic & Ground-Based & Bertin  \\ 
  21 & All-now & Space-Based & Avestruz, N. Li \& Lightman   \\ 
  22 & Manchester-NA2 & Ground-Based & Jackson \& Tagore  \\ 
  23 & YattaLensLite & Ground-Based & Sonnenfeld   \\ 
  24 & Kapteyn Resnet & Space-Based & Petrillo, Tortora, Verdoes Kleijn, Koopmans \& Vernardos \\ 
   \hline
\end{tabular}
\caption{Entries to the challenges.}
\label{table:entries}
\end{table*}

@@

Arcfinder \citep{2006astro.ph..6757A,2007A&A...461..813C,2012ApJ...749...38M} illustrated in Figure \ref{fig:Cabanac}, is a fast linear method that computes a pixel wise elongation parameter (ratio of first-order moments in a n-pix window oriented in proper reference frame) for all pixels of mexican-hat-smoothed FITS images. Arcfinder then extracts contiguous pixels above a given background and computes the candidate arc's length, width, area, radius of curvature and peak surface brightness. A final thresholding is set to maximize purity over completeness on a few typical arcs of the dataset.
For the current SL challenge, arcfinder was tuned to detect long and narrow arcs, and was optimized on a subset of 1000 simulated images with a grid covering a range of elongation windows and arc areas.  A python wrapper allows users to change parameters in a flexible way and run the arcfinder C code as a linux  command line. Arcfinder took a couple of hours to run on the entire dataset with some overheads due to the dataset format. The code is publicly available at https://github.com/rcabanac/arcfinder.

@@

For a binary classification problem such as the one involved in strong lens detection, the training is performed by optimizing the weights of the model as to minimize the \textit{binary cross-entropy} objective:
\begin{equation}
	- \sum_{n=1}^{N} y[n] \log \hat{y}[n] + (1 - y[n])\log(1 - \hat{y}[n]) \;,
\end{equation}
where $N$ is the number of training instances, $y \in \{0,1\}$ is the true class of the image and $\hat{y} \in [0,1]$ is the class probability predicted by the model. This optimization is usually performed by a Stochastic Gradient Descent (SGD) algorithm or its variants (e.g. ADAM \citep{Kingma_2014}, Adagrad \citep{duchi12adaptive}, RMSprop \citep{Tieleman12RMSProp}, or accelerated gradients \citep{nesterov83method}). SGD updates the model iteratively by taking small gradient steps over randomly selected subsamples of the training set (so called \textit{mini-batches}). All the CNN-based methods presented in this work rely on the ADAM optimisation algorithm, which also uses past gradients from previous iterations to adaptively estimate lower-order moments. Empirically it has been found that in many problems ADAM converges faster than SGD \citep{Ruder16overview}.

To avoid overfitting in neural networks, we used the Dropout regularisation technique \citep{2012arXiv1207.0580H,JMLR:v15:srivastava14a}. During training, we randomly drop a certain percent of neurones and their connections from the neural
network. This regularisation technique reduces overfitting  by preventing complex co-adaptations of neurones on training data.

Training multi-layer neural networks with gradient descent based approaches can be very challenging. One of the main reasons behind this is the effect of \textit{vanishing gradients}: it has been empirically observed that in many multi-layer neural networks the gradients in higher-level layers often become too small to be effective in gradient descent based optimisation. Another difficulty is that the distribution of each layers inputs changes during training as the parameters of the previous layers change. These issues make difficult to find the best learning rates.
